---

title: "Fraud Detection Analysis"
author: "Fraud Detection Internship Project"
date: "`r Sys.Date()`"
output:
html_document:
toc: true
toc_depth: 2
------------

# Overview

This report presents the Phase 11 analytical and visualization outputs
for the Fraud Detection System internship project.

The analysis uses compact aggregated datasets generated from the
processed PaySim dataset.

The project follows the internship methodology:

**Ask → Prepare → Process → Analyze → Share → Act**

This report contributes to the **Share** stage by communicating
fraud-related findings through tables and visualizations.

# Data Summary

```{r setup, message=FALSE, warning=FALSE}

library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)

data_dir <- file.path("..", "data", "visualization")

summary_data <- read_csv(
  file.path(data_dir, "fraud_summary.csv"),
  show_col_types = FALSE
)

by_type <- read_csv(
  file.path(data_dir, "fraud_by_type.csv"),
  show_col_types = FALSE
)

by_step <- read_csv(
  file.path(data_dir, "fraud_by_step.csv"),
  show_col_types = FALSE
)

by_amount <- read_csv(
  file.path(data_dir, "fraud_by_amount.csv"),
  show_col_types = FALSE
)

model_performance <- read_csv(
  file.path(data_dir, "model_performance.csv"),
  show_col_types = FALSE
)

summary_data
```

# Fraud by Transaction Type

The transaction-type analysis compares the number of transactions and
fraud rate associated with each transaction category.

```{r fraud-by-type-table}

by_type %>%
  arrange(desc(fraud_rate_percent))
```

```{r fraud-by-type-plot}

ggplot(
  by_type,
  aes(
    x = reorder(transaction_type, fraud_rate_percent),
    y = fraud_rate_percent
  )
) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Fraud Rate by Transaction Type",
    x = "Transaction Type",
    y = "Fraud Rate (%)"
  ) +
  theme_minimal()
```

# Fraud Across Simulation Steps

The PaySim dataset contains simulation steps that can be used to
investigate how fraud activity changes throughout the simulated period.

```{r fraud-by-step-plot}

ggplot(
  by_step,
  aes(
    x = step,
    y = fraud_transactions
  )
) +
  geom_line() +
  labs(
    title = "Fraud Transactions Across Simulation Steps",
    x = "Simulation Step",
    y = "Fraud Transactions"
  ) +
  theme_minimal()
```

# Transaction Amount Analysis

Transaction amounts are grouped into ranges so that the distribution
can be communicated without exposing the complete transaction-level
dataset.

```{r fraud-by-amount-table}

by_amount %>%
  arrange(amount_order)
```

```{r fraud-by-amount-plot}

ggplot(
  by_amount,
  aes(
    x = reorder(amount_bin, amount_order),
    y = fraud_rate_percent
  )
) +
  geom_col() +
  labs(
    title = "Fraud Rate by Transaction Amount Range",
    x = "Transaction Amount Range",
    y = "Fraud Rate (%)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  )
```

# Model Comparison

The Phase 8 model comparison provides the model-performance evidence
used in this visualization stage.

```{r model-performance-table}

model_performance
```

```{r model-performance-preparation}

model_long <- model_performance %>%
  select(
    model,
    precision,
    recall,
    f1_score,
    roc_auc
  ) %>%
  pivot_longer(
    cols = c(
      precision,
      recall,
      f1_score,
      roc_auc
    ),
    names_to = "metric",
    values_to = "score"
  )

model_long
```

```{r model-performance-plot}

ggplot(
  model_long,
  aes(
    x = model,
    y = score,
    fill = metric
  )
) +
  geom_col(position = "dodge") +
  labs(
    title = "Fraud Detection Model Performance",
    x = "Model",
    y = "Score",
    fill = "Metric"
  ) +
  ylim(0, 1.05) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(
      angle = 20,
      hjust = 1
    )
  )
```

# Findings

The visualization stage focuses on communicating the analytical
findings rather than creating additional machine-learning models.

The following areas are emphasized:

1. Overall fraud prevalence.
2. Differences in fraud rates between transaction types.
3. Changes in fraud activity across simulation steps.
4. Differences in fraud rates across transaction amount ranges.
5. Comparative model performance using precision, recall, F1-score,
   and ROC-AUC.

The Phase 8 model comparison identified Random Forest as the strongest
overall supervised approach in the evaluated results.

The anomaly-detection models remain important because they demonstrate
alternative approaches to fraud detection.

# Limitations

The PaySim dataset is synthetic and therefore the visual findings
should not be interpreted as production banking behaviour.

The visualizations communicate patterns in the project dataset and do
not establish that the same patterns would occur in a real financial
institution.

Model predictions identify transactions that may warrant further
investigation. They are not proof that a transaction or customer is
fraudulent.

# Conclusion

Phase 11 converts analytical results into visual and narrative outputs
for non-technical stakeholders.

The outputs include Python visualizations, R visualizations, an
R Markdown report, and an interactive Tableau dashboard.
